---
title: "Prediction Assignment"
author: "Patricialrt"
date: "24 January 2016"
output: html_document
---
```{r, echo=FALSE}
require(ggplot2)
require(caret)
require(lattice)
require(randomForest)
require(rpart) 
require(rpart.plot)
require(RColorBrewer)
#require(rattle) R crashes when trying to open rattle....
require(gbm)
set.seed(1234)
```
**Background**

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. 
In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har

**What you should submit**

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases. 

1. Your submission should consist of a link to a Github repo with your R markdown and compiled HTML file describing your analysis. Please constrain the text of the writeup to < 2000 words and the number of figures to be less than 5. It will make it easier for the graders if you submit a repo with a gh-pages branch so the HTML page can be viewed online (and you always want to make it easy on graders :-).
2. You should also apply your machine learning algorithm to the 20 test cases available in the test data above. Please submit your predictions in appropriate format to the programming assignment for automated grading. See the programming assignment for additional details.

```{r}
#Read the data
require(data.table)
trainUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
training <- read.csv(url(trainUrl), na.strings=c("NA","#DIV/0!",""))
testing <- read.csv(url(testUrl), na.strings=c("NA","#DIV/0!",""))
```


```{r}
#Clean Training data set
#Remove columns with more than 95% of NA or "" values
training <- training[, 6:dim(training)[2]]
treshold <- dim(training)[1] * 0.95
goodColumns <- !apply(training, 2, function(x) sum(is.na(x)) > treshold  || sum(x=="") > treshold)
training <- training[, goodColumns]
#Cleaning NearZeroVariance Variables Run this code to view possible NZV Variables:
badColumns <- nearZeroVar(training, saveMetrics = TRUE)
training <- training[, badColumns$nzv==FALSE]
training$classe = factor(training$classe)
```


```{r}
#Clean Testing data set
#Remove columns with more than 95% of NA or "" values
testing <- testing[, 6:dim(testing)[2]]
treshold <- dim(testing)[1] * 0.95
goodColumns <- !apply(testing, 2, function(x) sum(is.na(x)) > treshold  || sum(x=="") > treshold)
testing <- testing[, goodColumns]
#Cleaning NearZeroVariance Variables Run this code to view possible NZV Variables:
badColumns <- nearZeroVar(testing, saveMetrics = TRUE)
testing$classe = NA
testing <- testing[, badColumns$nzv==FALSE]
dim(testing)
```

```{r}
#Partition rows into training and crossvalidation, 60% for training:
require(caret)
inTrain <- createDataPartition(training$classe, p = 0.6)[[1]]
crossv <- training[-inTrain,]
training <- training[ inTrain,]
dim(training)
```


```{r}
#Train different models
require(caret)
modFit1 <- train(training$classe ~.,training,method="rpart") 
modFit1
results1 <- modFit1$results
round(max(results1$Accuracy),4)*100

ctrl   <- trainControl(method = "cv", number = 4, allowParallel = TRUE)
modFit2 <- train(training$classe ~.,training,method="rf",prof=TRUE, trControl = ctrl)
modFit2
results2 <- modFit2$results
round(max(results2$Accuracy),4)*100
# This method (rf) has much higher accuraccy as the first (rpart)

#Predicting in-sample errors
pred1 <- predict(modFit1, crossv)
pred2 <- predict(modFit2, crossv)
```

```{r}
#Use confusion matrices to test results/ cross-validation
confusionMatrix(pred1, crossv$classe)
confusionMatrix(pred2, crossv$classe)
```


```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(pred2)
```





